---
title: "Community event-based surveillance weekly report, Cox's Bazar"
author: "Cox's Bazar MSF-OCA Epidemiology Team" 
output:
  word_document:
    reference_docx: template_episitrep.docx
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---


```{r setup, include=FALSE, results='hide', message=FALSE, warning=FALSE}
## hide all code chunks in the output, but show errors
knitr::opts_chunk$set(echo = FALSE,       # hide all code chunks in output
                      error = TRUE,       # show errors if they appear, but don't stop
                      fig.width = 6*1.25, # Figure width
                      fig.height = 6,      # Figure height
                      warning = FALSE,
                      message = FALSE
                     )



## set default NA to - in output, define figure width/height
options(knitr.kable.NA = "-")



## Installing required packages for this template
required_packages <- c("knitr",       # create output docs
                       "here",        # find your files
                       "dplyr",       # clean/shape data
                       # "forcats",     # clean/shape data
                       "stringr",     # clean text
                       "rio",         # read in data
                       "ggplot2",     # create plots and charts
                       "patchwork",   # combine plots in one
                       "epikit",      # create categories from numerical variable
                       # "sitrep",      # MSF field epi functions
                       # "linelist",    # Functions for cleaning/standardising data/dates
                       # "matchmaker",  # dictionary-based standardization of variables
                       "incidence",   # create epicurves
                       # "aweek",       # define epi weeks
                       "sf",          # encode spatial vector data
                       "ggspatial",   # plot maps
                       "classInt",    # specifying breaks for maps
                       "tsibble",     # time series data
                       # "slider",      # time series data
                       "tidyr",       # long/long adjustements to data
                       "gt",          # make nice tables
                       "gtsummary",   # make nice tables
                       "viridis",     # additional palette to choose from
                       "data.table",  # for taking last and first values from dataframes
                       "tinytex")     # for making PDF

for (pkg in required_packages) {
  # install packages if not already present
  if (!pkg %in% rownames(installed.packages())) {
    install.packages(pkg)
  }
  
  # load packages to this current session 
  library(pkg, character.only = TRUE)
}



## Set default options for plots and charts

## set default text size to 16 for plots
## give classic black/white axes for plots
ggplot2::theme_set(theme_classic(base_size = 18))

## sets the theme in ggplot for epicurves
epicurve_theme <- theme(
  axis.text.x = element_text(angle = 90, hjust = 1, vjust = 1), 
  legend.title = element_blank(),
  panel.grid.major.x = element_line(color = "grey60", linetype = 3),
  panel.grid.major.y = element_line(color = "grey60", linetype = 3))
```


<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// define_current_week \\\
--------------------------------------------------------------------------------

You need to set the week you want to report on. Generally, this is the previous
week. Put it below.

aweek::set_week_start will define the beginning of the week. The standard is
Monday.
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->

<!-- **This section will need to be updated weekly** -->
<!-- ##[EDIT] -->
```{r define_current_week}

## set current week and start date on Sunday
reporting_week <- yearweek(Sys.Date(),week_start = 7) -1

## Set previous week
last_week <- reporting_week - 1

# Set two weeks ago
two_weeks_ago <- reporting_week - 2

# Set three weeks ago
three_weeks_ago <- reporting_week - 3

```
Data as reported by `r reporting_week`.



<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// read_nonDHIS_data \\\
---------------------------------------------------------------------------------->

```{r read_nonDHIS_data, warning = FALSE, message = FALSE}

## EBS data 
ebs_raw  <- rio::import(here::here("2021", "2_data", "EBS form.xlsx")) %>% 
  ## make all colnames lower case
  janitor::clean_names() 

```


<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// Data cleaning \\\
---------------------------------------------------------------------------------->

<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
This part of the script will create and clean variables in your data.

All your cleaning and variable creation should happen in these chunks.
That way, in case something goes wrong, you can push the small arrow at the top
of the chunk to re-run all the code chunks up to the current one.

The chunks are:
- standardise_dates -- will set up and clean dates.
- create_vars       -- creates variables based on other variables

You must adapt this section according to your data!
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->


<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// standardise_dates \\\
--------------------------------------------------------------------------------

This chunk will help you set up and clean your date variables.
Also removes any empty columns
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->

```{r standardise_dates}

## use the guess_dates() function to make a first pass at date variables.
ebs_cleaned <- ebs_raw %>%
  # mutate_at(c("chv_signals_date_case_reported",
  #             "verification_date_of_assessment"),
  #           linelist::guess_dates,
  #           error_tolerance = 0.5) %>% 
  ## remove empty columns
  janitor::remove_empty("cols")

```


<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// create_vars \\\
--------------------------------------------------------------------------------

This chunk will help you construct new variables from other variables. 
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->
```{r create_vars}

## create epiweeks for the signals and assessments
ebs_cleaned <- ebs_cleaned %>% 
  ## create an epi week signal variable
  mutate(epiweek_signal = yearweek(chv_signals_date_case_reported, 
                                           week_start = 7)) %>% 
  ## create an epi week verification signal
  mutate(epiweek_invest = yearweek(verification_date_of_assessment, 
                                                  week_start = 7)) %>% 
  ## create an epi week risk assessment
  mutate(epiweek_ra = yearweek(date_of_risk_assessment, 
                                                  week_start = 7))


## recode the event variable
ebs_cleaned <- ebs_cleaned %>% 
  mutate(chv_signals_event = case_when(
    chv_signals_event == "1" ~ "Illness",
    chv_signals_event == "8" ~ "Death",
    chv_signals_event == "10" ~ "Unusual health event",
    TRUE ~ NA_character_
  ))


## make a combined event variable that takes into consideration the details
ebs_cleaned <- ebs_cleaned %>% 
  mutate(total_event  = case_when(
    chv_signals_event == "Illness" & chv_signals_eventdescription == "AJS" ~ "Illness (regular outbreak-prone)",
    chv_signals_event == "Illness" & chv_signals_eventdescription == "AWD" ~ "Illness (regular outbreak-prone)",
    chv_signals_event == "Illness" & chv_signals_eventdescription == "AFP" ~ "Illness (regular outbreak-prone)",
    chv_signals_event == "Illness" & chv_signals_eventdescription == "Measles" ~ "Illness (regular outbreak-prone)",
    chv_signals_event == "Illness" & chv_signals_eventdescription == "Meningitis" ~ "Illness (regular outbreak-prone)",
    chv_signals_event == "Illness" & chv_signals_eventdescription == "Diphtheria" ~ "Illness (regular outbreak-prone)",
    chv_signals_event == "Illness" & chv_signals_eventdescription == "ARTI/Suspected Covid" ~  "Suspected COVID",
    chv_signals_eventdescription == "Skin disease" ~ "Skin disease",
    chv_signals_event == "Death" & chv_signals_eventdescription == "ARTI/Suspected Covid" ~ "Suspected COVID death",
    chv_signals_event == "Death" ~ "Other death",
    chv_signals_event  == "Livestock" ~ "Unusual health event",
    chv_signals_eventdescription == "livestock" ~ "Unusual health event",
    chv_signals_event == "Unusual health event" ~ "Unusual health event",
    TRUE ~ NA_character_
  ))


## make a detail event variable
ebs_cleaned <- ebs_cleaned %>% 
  mutate(total_event_detail = case_when(
    chv_signals_event == "Illness" & chv_signals_eventdescription == "AJS" ~ "AJS",
    chv_signals_event == "Illness" & chv_signals_eventdescription == "AWD" ~ "AWD",
    chv_signals_event == "Illness" & chv_signals_eventdescription == "AFP" ~ "AFP",
    chv_signals_event == "Illness" & chv_signals_eventdescription == "Measles" ~ "Measles",
    chv_signals_event == "Illness" & chv_signals_eventdescription == "Meningitis" ~ "Meningitis",
    chv_signals_event == "Illness" & chv_signals_eventdescription == "Diphtheria" ~ "Diphtheria",
    chv_signals_event == "Illness" & chv_signals_eventdescription == "ARTI/Suspected Covid" ~  "Suspected COVID",
    chv_signals_eventdescription == "Skin disease" ~ "Skin disease",
    chv_signals_event == "Death" & chv_signals_eventdescription == "ARTI/Suspected Covid" ~ "Suspected COVID death",
    chv_signals_event == "Death" ~ "Other death",
    chv_signals_eventdescription == "livestock" ~ "Unusual health event",
    chv_signals_event == "Unusual health event" ~ "Unusual health event",
    TRUE ~ NA_character_
  ))


## make a signal status variable
ebs_cleaned <- ebs_cleaned %>% 
  mutate(signal_status = case_when(
    verification_is_the_event_confirem == "yes" ~ "Verified",
    chv_signals_verification_req == "yes"  | 
      chv_signals_verification_req == "no" ~ "Investigated (not valid)",
    TRUE ~ "Not investigated"
  )) %>% 
  ## convert signal status to a factor and specify the levels
  mutate(signal_status = factor(signal_status,
                                levels = c("Verified", 
                                           "Investigated (not valid)",
                                           "Not investigated")))


## make a camp variable to match with the shapefile
ebs_cleaned <- ebs_cleaned %>% 
  ## create a camp variable that adds "Camp " in front of the number
  mutate(site_name  = paste0("Camp ",chv_signals_camp)) %>% 
  ## replace lower case e to upper case
  mutate(site_name = str_replace(site_name, "e", "E")) %>% 
  ## replace lower case w to upper case
  mutate(site_name = str_replace(site_name, "w", "W"))

```

<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// make additional data subset \\\
--------------------------------------------------------------------------------

This chunk will help you construct a data subset for mapping purposes 
- AWD
- AJS
- Diphtheria
- Suspected COVID-19
- Measles
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->

```{r functions}

## create a function to extract count data for specific verified signals past week
count_verified_signals_past_week <- function(data, signal =
                                               c("AWD",
                                                 "AJS", 
                                                 "Measles",
                                                 "Meningitis",
                                                 "Diphtheria", 
                                                 "Suspected COVID"))
  {
  data %>% 
    ## specify time period to keep 
    filter(epiweek_signal == reporting_week ) %>% 
    ## only show the verified signals
    filter(signal_status == "Verified") %>% 
    ## filter on total_event detail
    filter(total_event_detail == signal) %>% 
    ## select variables of interest
    select(site_name) %>% 
    group_by_all() %>% 
    tally()
}  

## create a function to extract count data for specific verified signals past 4weeks
count_verified_signals_past_4weeks <- function(data, signal =
                                               c("AWD",
                                                 "AJS", 
                                                 "Diphtheria", 
                                                 "Measles",
                                                 "Meningitis",
                                                 "Suspected COVID"))
  {
  data %>% 
    ## specify time period to keep 
    filter(epiweek_signal >= three_weeks_ago & epiweek_signal <= reporting_week ) %>% 
    ## only show the verified signals
    filter(signal_status == "Verified") %>% 
    ## filter on total_event detail
    filter(total_event_detail == signal) %>% 
    ## select variables of interest
    select(site_name) %>% 
    group_by_all() %>% 
    tally()
    }                                                


```

<!-- ## Automated text for reporting paragraph -->

```{r reporting}

## count of weekly signals
total_signals_received <- ebs_cleaned %>% 
  ## choose only the signals from the reporting week
  filter(epiweek_signal == reporting_week) 
  

## Count of weekly signals investigated
total_signals_investigated <- ebs_cleaned %>% 
  ## choose only the signals from the reporting week
  filter(epiweek_signal == reporting_week) %>% 
  ## choose only the signals that were verified or investigated and not validated
  filter(signal_status %in% c("Verified", "Investigated (not valid)"))

## count of weekly signals verified
verified_signals_investigated <- ebs_cleaned %>% 
  ## choose only the signals from the reporting week
  filter(epiweek_signal == reporting_week) %>% 
  ## choose only the signals that were verified
  filter(signal_status == "Verified")

## count and proportion of investigated signals by who they were received from
identification_investigated_signals <- ebs_cleaned %>% 
  ## choose only the signals from the reporting week
  filter(epiweek_signal == reporting_week) %>% 
  ## choose only the signals that were verified or investigated and not validated
  filter(signal_status %in% c("Verified", "Investigated (not valid)")) %>% 
  ## select the who reported the data variable
  select(chv_signals_cases_reported_by_who) %>% 
  group_by_all() %>% 
  ## count the number of signals received per category
  summarise(n = n()) %>% 
  ungroup() %>% 
  mutate(proportion = round((n/sum(n, na.rm = TRUE))*100, digits = 0))


          
## Proportion of investigated signals received by CHWs
prop_chv <- identification_investigated_signals %>% 
  ## only select the number and percent received by CHVs
  filter(chv_signals_cases_reported_by_who == "msf_worker" |
          chv_signals_cases_reported_by_who == "hp_chv" ) %>% 
  ## obtain the %
  pull(proportion)

## Proportion of investigated signals received by EAT
prop_eat <- identification_investigated_signals %>% 
  ## only select the number and percent received by CHVs
  filter(chv_signals_cases_reported_by_who == "eat") %>% 
  ## obtain the %
  pull(proportion)
  

## Calculate the four week average
four_week_signal_average  <- ebs_cleaned %>% 
  ## keep last four weeks
  filter(epiweek_signal >= three_weeks_ago & epiweek_signal <= reporting_week) %>% 
  ## choose only the signals that were verified or investigated and not validated
  filter(signal_status %in% c("Verified", "Investigated (not valid)")) %>% 
  ## select variable of interest
  select(epiweek_signal, signal_status) %>% 
  group_by(epiweek_signal) %>% 
  tally() %>% 
  ## average of past four weeks
  summarise(average = as.integer(mean(n, na.rm = TRUE)))
  
trend_investigated_signals <- if (four_week_signal_average > nrow(total_signals_investigated)) {
  "a decreasing"
} else if (four_week_signal_average < nrow(total_signals_investigated)) {
  "increasing"
} else {"no change"}

## number of referrals by event type
number_referred <- ebs_cleaned %>%
  ## choose only the signals from the reporting week
  filter(epiweek_invest == reporting_week) %>%
  select(total_event_detail, verification_number_of_people_ref_arest_health_faciliy) %>%
  filter(!is.na(verification_number_of_people_ref_arest_health_faciliy)) %>%
  ## sum of cases in week
  group_by(total_event_detail) %>%
  summarise(Referred=sum(verification_number_of_people_ref_arest_health_faciliy))

## number of referrals in total
number_referred_total <- number_referred %>%
  summarise(Referred=sum(Referred))

##number of referrals that required follow up
number_referred_fu <- ebs_cleaned %>%
    ## choose only the signals from the reporting week
  filter(epiweek_invest == reporting_week) %>%
  select(total_event_detail, verification_referal_fu) %>%
  filter(!is.na(verification_referal_fu)) %>%
  ## sum of cases in week
  group_by(total_event_detail) %>%
  summarise(Referred_FU=sum(verification_referal_fu))

## number of fu referrals in total
number_referred_fu_total <- number_referred_fu %>%
  summarise(Referred_FU=sum(Referred_FU))

## number of risk assessments done
risk_as <-ebs_cleaned %>%
    ## choose only the risk assessments from the reporting week
  filter(epiweek_ra == reporting_week) 

##number of risk assessments that require response
risk_as_resp <-ebs_cleaned %>%
    ## choose only the risk assessments from the reporting week
  filter(epiweek_ra == reporting_week) %>%
    ## chose only the risk assessments categorised as minor- severe
  filter(what_was_the_risk_level =="minor"|what_was_the_risk_level =="moderate"|what_was_the_risk_level =="major"|what_was_the_risk_level =="severe")
  
```

<!-- ## Automated text for regular outbreak reporting -->


```{r regular_outbreak}
## count of illness signals investigated
illness_signals_investigated <- ebs_cleaned %>% 
  ## only keep the signals from the reporting week
  filter(epiweek_signal == reporting_week) %>% 
   ## choose only the signals that were verified or investigated and not validated
  filter(signal_status %in% c("Verified", "Investigated (not valid)")) %>% 
  ## only choose those that correspond to illnesses
  filter(total_event == "Illness (regular outbreak-prone)")


## Calculate the four week average of illness signals
four_week_illness_signal_average  <- ebs_cleaned %>% 
  ## keep last four weeks
  filter(epiweek_signal >= three_weeks_ago & epiweek_signal <= reporting_week) %>% 
  ## choose only the signals that were verified or investigated and not validated
  filter(signal_status %in% c("Verified", "Investigated (not valid)")) %>% 
  ## only keep the illness signals
  filter(total_event == "Illness (regular outbreak-prone)") %>% 
  ## select variable of interest
  select(epiweek_signal, signal_status) %>% 
  group_by(epiweek_signal) %>% 
  tally() %>% 
  ## average of past four weeks
  summarise(average = as.integer(mean(n, na.rm = TRUE)))

## Check compared to the average of the past 4 weeks if current week is higher?
trend_investigated_illness_signals <- if (four_week_illness_signal_average > nrow(illness_signals_investigated)) {
  "lower"
} else if (four_week_illness_signal_average < nrow(illness_signals_investigated)) {
  "higher"
} else {"the same"}

```

<!-- ## Automated text for suspected COVID-19 cases -->

```{r suspected_c19}

## count of covid19 signals investigated
covid19_signals_investigated <- ebs_cleaned %>% 
  ## only keep the signals from the reporting week
  filter(epiweek_signal == reporting_week) %>% 
   ## choose only the signals that were verified or investigated and not validated
  filter(signal_status %in% c("Verified", "Investigated (not valid)")) %>% 
  ## only choose those that correspond to illnesses
  filter(total_event == "Suspected COVID")


## Calculate the four week average of covid19 signals
four_week_covid19_signal_average  <- ebs_cleaned %>% 
  ## keep last four weeks
  filter(epiweek_signal >= three_weeks_ago & epiweek_signal <= reporting_week) %>% 
  ## choose only the signals that were verified or investigated and not validated
  filter(signal_status %in% c("Verified", "Investigated (not valid)")) %>% 
   ## only choose those that correspond to illnesses
  filter(total_event == "Suspected COVID") %>% 
  ## select variable of interest
  select(epiweek_signal, signal_status) %>% 
  group_by(epiweek_signal) %>% 
  tally() %>% 
  ## average of past four weeks
  summarise(average = as.integer(mean(n, na.rm = TRUE)))

## Check compared to the average of the past 4 weeks if current week is higher?
trend_investigated_covid_signals <- if (four_week_covid19_signal_average > nrow(covid19_signals_investigated)) {
  "a decrease"
} else if (four_week_covid19_signal_average < nrow(covid19_signals_investigated)) {
  "an increase"
} else {"no change"}

```

<!-- ## Automated text for unusual health events -->

```{r unusual_health_events}

## count of unusual event signals investigated
unusual_event_signals_investigated <- ebs_cleaned %>% 
  ## only keep the signals from the reporting week
  filter(epiweek_signal == reporting_week) %>% 
   ## choose only the signals that were verified or investigated and not validated
  filter(signal_status %in% c("Verified", "Investigated (not valid)")) %>% 
  ## only choose those that correspond to illnesses
  filter(total_event == "Unusual health event")

## count of verified unusual health event verified
verified_unusual_event <- unusual_event_signals_investigated %>% 
  ## only keep those that were verified
  filter(signal_status == "Verified")

```


<!-- ## Automated text for deaths -->

```{r deaths}

## count of deaths investigated
death_signals_investigated <- ebs_cleaned %>% 
  ## only keep the signals from the reporting week
  filter(epiweek_signal == reporting_week) %>% 
   ## choose only the signals that were verified or investigated and not validated
  filter(signal_status %in% c("Verified", "Investigated (not valid)")) %>% 
  ## only choose those that correspond to illnesses
  filter(total_event %in% c("Other death", "Suspected COVID death"))


## count of suspected covid19 deaths
covid_deaths <- death_signals_investigated %>% 
  ## only count the covid-19 deaths
  filter(total_event == "Suspected COVID death")

## count of verified suspected covid19 deaths
verified_covid_deaths <- death_signals_investigated %>% 
  ## only count the covid-19 deaths
  filter(total_event == "Suspected COVID death") %>% 
  ## only keep the verified signals
  filter(signal_status == "Verified")

```
<!-- ## Automated text for Thresholds and recommendations -->

```{r thresholda}

thresholda <- ebs_cleaned %>%
  ## only keep signals reported during the reporting week 
  filter(epiweek_signal == reporting_week) %>% 
  ## only keep the verified signals
  filter(signal_status == "Verified")  %>%
  ## only keep signals that require a response
  filter(verification_do_you_think_response_is_requi == "yes")  %>%
  ## only keep signals where no other actor is responding
  filter(verification_is_there_any_other_organizatio == "no")  %>%
  ## count number of signals per event type, per camp
  count(chv_signals_camp, total_event_detail, sort=TRUE)

thresholda <- thresholda[1,]

```

```{r thresholdb}
 thresholdb <-ebs_cleaned %>%
  ## only keep signals reported during the reporting week 
  filter(epiweek_signal == reporting_week) %>% 
  ## only keep the verified signals
  filter(signal_status == "Verified")  %>%
  ## only keep signals that require a response
  filter(verification_do_you_think_response_is_requi == "yes")  %>%
  ## only keep signals where no other actor is responding
  filter(verification_is_there_any_other_organizatio == "no")  %>%
  group_by(chv_signals_camp) %>%
  ## by camp, keep the max number of people affected
  filter(verification_how_many_people == max(verification_how_many_people) ) %>% 
  ## Select the camp, detail and the number of people affected
              select(chv_signals_camp,total_event_detail,    verification_how_many_people)

## arrange the datbase in order from big to small in terms of # affected
thresholdb <- thresholdb %>%
  arrange(desc(verification_how_many_people))
## Extract the event with the greatest number of people affected
thresholdb <- thresholdb[1,]
 


```

```{r thresholdc}

thresholdc  <- ebs_cleaned %>% 
  ## keep last three weeks
  filter(epiweek_signal >= two_weeks_ago & epiweek_signal <= reporting_week) %>% 
  ## choose only the signals that were verified 
  filter(signal_status == "Verified")  %>%
  ## only keep signals that require a response
  filter(verification_do_you_think_response_is_requi == "yes")  %>%
  ## only keep signals where no other actor is responding
  filter(verification_is_there_any_other_organizatio == "no")  %>%
  ##drop signals that are not categorised
  drop_na(total_event_detail) %>%
  ## select variable of interest
  select( epiweek_signal, total_event_detail, verification_how_many_people) %>% 
  group_by(total_event_detail, epiweek_signal) %>% 
  ## sum of cases in week
  summarise(n = sum(verification_how_many_people))

thresholdc <- thresholdc %>%
  ## make a wide dataseat
  spread(epiweek_signal, n, fill=0) %>% 
  ## rename to show that it's signals
  dplyr::rename("Signals" = total_event_detail) 

## replace "NA" with 0
thresholdc[is.na(thresholdc)] <- 0

## Create a variable that shows if there has been a consecutive increase in the past 3 weeks
 thresholdc$increase <- ifelse(
    thresholdc[[as.character(two_weeks_ago)]] < thresholdc[[as.character(last_week)]] & 
      thresholdc[[as.character(last_week)]] < thresholdc[[as.character(reporting_week)]],
    TRUE, 
    FALSE)  
 

thresholdc_increase <- thresholdc %>%
  filter(increase==TRUE) %>%
  ## arrange the datbase in order from big to small in terms of # affected
  arrange(desc(reporting_week)) %>%
  ## drop livestock, unusual events and other deaths from this- since these categories are formed of many subcategories that need to be further reviewed before recommending a risk assessment
  filter( Signals!= "Other death" & Signals!="Unusual health event")
## Extract the event with the greatest number of people affected
thresholdc_increase <- thresholdc_increase[1,]


```

# Summary  
This community event-based surveillance (EBS) report covers topics of cluster of illness, cluster of death or increased death, unusual health events, and livestock diseases in Cox's Bazar in `r reporting_week` 

## Reporting   
  - During `r reporting_week`, there were a total of `r  nrow(total_signals_received)` signals received of which `r nrow(total_signals_investigated)` signals were investigated by the Epidemiology Alert Team (EAT) and  `r nrow(verified_signals_investigated)` verified.  `r prop_chv`% of signals were identified by HP Community Health Volunteers (CHVs) or directly by the Epi team (`r prop_eat`%). There is `r trend_investigated_signals` trend in investigated signals compared to the average of the previous four weeks (`r four_week_signal_average %>% pull(average)`). The teams made `r number_referred_total %>% pull(Referred)` verbal referrals to the person's nearest health facility, including `r number_referred_fu_total %>% pull(Referred_FU)` people who required follow up at MSF OCA facility. `r nrow(risk_as)` risk assessments were done and `r nrow(risk_as_resp)` events were classified as more than minimal risk  (minor-severe) requiring additional response and resources- see attachments.
  
## Regular outbreak reporting  

  - `r nrow(illness_signals_investigated)` outbreak-prone illness signals were investigated in `r reporting_week`, of which `r fmt_count(illness_signals_investigated, total_event_detail == "AFP")` AFP,  `r fmt_count(illness_signals_investigated, total_event_detail == "AJS")` AJS, `r fmt_count(illness_signals_investigated, total_event_detail == "AWD")` AWD, `r fmt_count(illness_signals_investigated, total_event_detail == "Diphtheria")` Diphtheria,  `r fmt_count(illness_signals_investigated, total_event_detail == "Measles")` measles and `r fmt_count(illness_signals_investigated, total_event_detail == "Menigitis")` meningitis. `r fmt_count(illness_signals_investigated, signal_status == "Verified")` of the illness signals were verified by the EAT. The total illness signals investigated in `r reporting_week` is `r trend_investigated_illness_signals` compared to the four week average of `r four_week_illness_signal_average %>% pull(average) `.

## Suspected COVID-19 cases  
   - During `r reporting_week`,`r nrow(covid19_signals_investigated)` signals for suspected COVID-19 were investigated and `r fmt_count(covid19_signals_investigated, verification_how_many_people >= 2)` of them were clusters of two or more cases. There is `r trend_investigated_covid_signals ` in `r reporting_week` compared to the average of the past four weeks (`r four_week_covid19_signal_average %>% pull(average)`)
   
## Unusual health events  
  - During `r reporting_week`, there was a total of `r nrow(unusual_event_signals_investigated)` unusual health events investigated. Among them `r nrow(verified_unusual_event)` (`r round((nrow(verified_unusual_event)/nrow(unusual_event_signals_investigated))*100, digits = 0)`%) were verified. 


## Deaths  
  - There were a total of `r nrow(death_signals_investigated)` deaths investigated, of which `r fmt_count(death_signals_investigated, verification_how_many_people == 1)` single deaths and `r fmt_count(death_signals_investigated, verification_how_many_people >= 2)` clusters of deaths investigated. `r fmt_count(death_signals_investigated, total_event == "Suspected COVID death")` ARTI/suspected COVID-19 death signals were investigated and `r nrow(verified_covid_deaths)` were verified.
  
# Recommendations  
  - Risk assessment for `r thresholda %>% pull(total_event_detail)` in camp `r thresholda %>% pull(chv_signals_camp)` since there  were `r thresholda %>% pull(n)` clusters in this camp in the reporting week
  
  - Risk assessment for `r thresholdb %>% pull(total_event_detail)` in camp `r thresholdb %>% pull(chv_signals_camp)` since there was 1 large cluster that affected `r thresholdb %>% pull(verification_how_many_people)` people. 
  
  - Risk assessment for `r thresholdc_increase %>% pull(Signals)` since there has been an increase in the number of people affected across the camps in the past three consecutive weeks and `r thresholdc_increase %>% pull(c(4))` people affected in the reporting week
  
<!-- ##Here add any other recommendations- including any facility based information that may be linked, or info from risk assessments  -->
<!-- ##[EDIT] -->
[INSERT ANY OTHER SUMMARY TEXT HERE]
  

\newpage



<!-- ## Table 1 -->
<!-- Create summary table for the week -->
## EBS activities in `r reporting_week`
<!-- ##[EDIT] -->
[INSERT SUMMARY TEXT HERE]



```{r summary_table}

ebs_cleaned %>% 
  ## only keep signals reported during the reporting week
  filter(epiweek_signal == reporting_week) %>% 
  ## select variables to use for table
  select("Signal type" = total_event, 
         signal_status) %>% 
  ## use tbl_summary from gtsummary package to make table
  tbl_summary(by = signal_status, 
              missing = "no") %>% 
  ## add total for the week
  add_overall() %>%
  bold_labels() %>%
  as_gt() %>%
  gt::gtsave(filename="summary.png", path = here::here("2021", "4_output", "archive"))
```

## Verified signals in `r reporting_week`
<!-- ##[EDIT] -->
[INSERT SUMMARY TEXT HERE]

<!-- ## Figure 1 - Number of verified signals per week -->
```{r plot_verified_signals, fig.height=5}

## make a dataset that only included verified signals
ebs_verified <- ebs_cleaned %>% 
  ## only  keep the verified signals
  filter(signal_status == "Verified") %>% 
  ## remove any NAs from total_event_detail
  filter(is.na(total_event_detail) == FALSE) %>% 
  ## only keep those with epiweek data
  filter(is.na(epiweek_signal) == FALSE) %>% 
  ## only include signals from reporting_week and before
  filter(epiweek_signal <= reporting_week)

## make a subset which is just for the plot
ebs_verified_plot <- ebs_verified %>% 
  ## only  keep outbreak-prone illnesses
  filter(total_event_detail == "AFP" | total_event_detail == "AJS"| total_event_detail == "AWD"| total_event_detail == "Diphtheria" | total_event_detail == "Measles"| total_event_detail == "Suspected COVID") %>%
  ## group by week and event detail
  group_by(epiweek_signal, total_event_detail) %>% 
  ## tally
  tally()
  ## clean epiweek for clearer labeling
ebs_verified_plot$epiweek_signal <-  substr(ebs_verified_plot$epiweek_signal, 7, 8)


# ##THIS IS THE OLD STACKED GRAPH SCRIPT
#   ## Make a plot
# ggplot(ebs_verified_plot, aes(x = epiweek_signal, y = n, 
#                          fill = total_event_detail)) +
#   ## stacked barplot
#   geom_col(position = "stack") +
#   ## Use the viridis colour choice
#   scale_fill_viridis(discrete = TRUE, option = "D" ) +
#   ## standardised aesthetic modifications to plot
#   theme_classic() +
#   ## add labels
#   labs(x = "Week",
#        y = "Number of signals",
#        fill = "Detail of verified signals",
#        title = "Verified outbreak-prone signals by epidemiological week")


## bar graph non stacked per outbreak illness
ggplot(ebs_verified_plot, aes(x = epiweek_signal, y = n, fill = total_event_detail)) +
  ## barplot
  geom_col(position="dodge") +
  geom_text(aes(label=n), vjust= -0.5, size=2.5) +
  ## Use the viridis colour choice
  scale_fill_viridis(discrete = TRUE, option = "D" ) +
  ggtitle("Outbreak-prone illness signals per epi week") +
  facet_wrap(~total_event_detail) +
  theme_minimal() +
  theme (text=element_text(size=10), axis.text.x= element_text(angle=90), legend.position= "none") +
   ## add labels
  labs(x = "Epi week",
       y = "Number of signals")
```



<!-- ## Map -->
<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// read_shapefiles \\\
--------------------------------------------------------------------------------

To create maps, you need to have a shapefile of the area. Often, the MSF GIS
unit can provide shapefiles.

Your shapefile can be a polygon or points. Polygons do not need to be contiguous.

The names of the polygons or points MUST match the names in your linelist.

Your coordinate reference system needs to be WGS84.
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->
```{r read_shapefiles, message=FALSE}

## read in shapefile
map <- read_sf(here::here("../../2_GIS/GIS/MSF_GIS_Folder/Camp shape file/Site_boundaries.shp")) %>% 
  ## change all variables to lower case
  janitor::clean_names()

## create a subset of map file with only the main camp area
map <- map %>% 
  filter(!site_name %in% c("Camp 21", "Camp 22", 
                           "Camp 23", "Camp 24",
                           "Camp 25", "Camp 26", 
                           "Camp 27", "Choukali",
                           "Nayapara RC"))
```


<!-- ## Base layer from OSM -->

```{r obtain_base_layer}

tiles <- cartography::getTiles(
  map,
  type = "OpenStreetMap",
  zoom = NULL,
  crop = TRUE,
  verbose = FALSE,
  apikey = NA,
  cachedir = FALSE,
  forceDownload = FALSE
)

```


<!-- ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
/// choropleth_maps \\\
--------------------------------------------------------------------------------

Once you have loaded your shapefile, you can map the case counts or attack rates.

Make sure you delete or comment out the section you are not using.
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ -->

<!-- ## Suspected COVID map in the past week -->

```{r choropleth_covid_pw_maps, message = FALSE, warning = FALSE}

## Join the past month dataset with map file
covid_past_week <- map %>% 
  ## join the count data from past month verified dataset
  left_join(count_verified_signals_past_week(ebs_cleaned, 
                                             "Suspected COVID"),
            by = "site_name") %>% 
  
    ## create levels for counts of covid
  mutate(covid_counts =  fac_from_num(n))%>% 
  ## remove the NA from covid_counts
  filter(is.na(covid_counts) == FALSE)
  


## Plot map of cases by block
covid_past_week_map <- ggplot() +
  ## add in the back ground tiles
  ggspatial::layer_spatial(tiles, interpolate = TRUE) +
  # shapefile as polygon
  geom_sf(data = covid_past_week, aes(fill = covid_counts)) + 
  # needed to avoid gridlines being drawn
  coord_sf(datum = NA) + 
  # add a scalebar
  annotation_scale() + 
  # choose palette colour for fill
  scale_fill_brewer(palette = "YlOrRd") +
  ## Add camp number as label
  geom_sf_text(data = covid_past_week, aes(label = site_name), colour = "black",
               size = 2.5) +
  theme_void() +
  theme(plot.title = element_text(size= 12)) +
  labs(fill = "No. of signals",
       captions = str_glue("Source: MSF data from {reporting_week}"),
       title = str_glue("Number of verified suspected COVID-19\n signals in Cox's Bazar, {reporting_week}"))
  



```

<!-- ## Suspected COVID map in the past week -->

```{r choropleth_covid_p4w_maps, message = FALSE, warning = FALSE}

## Join the past month dataset with map file
covid_past_4weeks <- map %>% 
  ## join the count data from past month verified dataset
  left_join(count_verified_signals_past_4weeks(ebs_cleaned, 
                                             "Suspected COVID"),
            by = "site_name") %>% 
    ## create levels for counts of covid
  mutate(covid_counts =  fac_from_num(n)) %>% 
  ## remove the NA from covid_counts
  filter(is.na(covid_counts) == FALSE)
  


## Plot map of cases by block
covid_past_4weeks_map <- ggplot() +
  ## add in the back ground tiles
  ggspatial::layer_spatial(tiles, interpolate = TRUE) +
  # shapefile as polygon
  geom_sf(data = covid_past_4weeks, aes(fill = covid_counts)) + 
  # needed to avoid gridlines being drawn
  coord_sf(datum = NA) + 
  # add a scalebar
  annotation_scale() + 
  # choose palette colour for fill
  scale_fill_brewer(palette = "YlOrRd") +
  ## Add camp number as label
  geom_sf_text(data = covid_past_week, aes(label = site_name), colour = "black",
               size = 2.5) +
  theme_void() +
  theme(plot.title = element_text(size= 12)) +
  labs(fill = "No. of signals",
       captions = str_glue("Source: MSF data from {three_weeks_ago}-
                        {reporting_week}"),
       title = str_glue("Number of verified suspected COVID-19\n signals in Cox's Bazar,\n {three_weeks_ago}-{reporting_week}"))
  


```

<!-- #### Combine covid maps -->
```{r combine_covid_maps, fig.height = 16, fig.height = 5}

covid_past_4weeks_map  + covid_past_week_map 

```


<!-- ## AWD past week -->
```{r choropleth_awd_pw_maps, message = FALSE, warning = FALSE}

## Join the past month dataset with map file
awd_past_week <- map %>% 
  ## join the count data from past month verified dataset
  left_join(count_verified_signals_past_week(ebs_cleaned, "AWD"), 
            by = "site_name") %>% 
  ## create levels for counts of awd
  mutate(awd_counts = fac_from_num(n)) %>% 
  ## remove the empty values
  filter(is.na(awd_counts) == FALSE)


## Plot map of cases by camp
awd_past_week_map <- ggplot() +
  ## add in the back ground tiles
  ggspatial::layer_spatial(tiles, interpolate = TRUE) +
  # shapefile as polygon
  geom_sf(data = awd_past_week, aes(fill = awd_counts)) + 
  # needed to avoid gridlines being drawn
  coord_sf(datum = NA) + 
  # add a scalebar
  annotation_scale() + 
  # choose palette colour for fill
  scale_fill_brewer(palette = "YlOrRd") +
  ## Add camp number as label
  geom_sf_text(data = awd_past_week, aes(label = site_name), colour = "black",
               size = 2.5) +
  theme_void() +
  theme(plot.title = element_text(size= 12)) +
  labs(fill = "No. of signals",
       captions = str_glue("Source: MSF data from {reporting_week}"),
       title = str_glue("Number of verified AWD signals in \nCox's Bazar,{reporting_week}"))
  
```

<!-- ## AWD past 4 weeks -->

```{r choropleth_awd_p4w_maps, message = FALSE, warning = FALSE}

## Join the past month dataset with map file
awd_past_4weeks <- map %>% 
  ## join the count data from past month verified dataset
  left_join(count_verified_signals_past_4weeks(ebs_cleaned, "AWD"), 
            by = "site_name") %>% 
  ## create levels for counts of awd
  mutate(awd_counts = fac_from_num(n)) %>% 
  ## remove the empty values
  filter(is.na(awd_counts) == FALSE)
         


## Plot map of cases by camp
awd_past_4weeks_map <- ggplot() +
  ## add in the back ground tiles
  ggspatial::layer_spatial(tiles, interpolate = TRUE) +
  # shapefile as polygon
  geom_sf(data = awd_past_4weeks, aes(fill = awd_counts)) + 
  # needed to avoid gridlines being drawn
  coord_sf(datum = NA) + 
  # add a scalebar
  annotation_scale() + 
  # choose palette colour for fill
  scale_fill_brewer(palette = "YlOrRd") +
  ## Add camp number as label
  geom_sf_text(data = awd_past_4weeks, aes(label = site_name), colour = "black",
               size = 2.5) +
  theme_void() +
  theme(plot.title = element_text(size= 12)) +
  labs(fill = "No. of signals",
       captions = str_glue("Source: MSF data from {three_weeks_ago} to {reporting_week}"),
       title = str_glue("Number of verified AWD signals in \nCox's Bazar, {three_weeks_ago}-{reporting_week}"))
  
```

<!-- #### Combine AWD maps -->
```{r combine_awd_maps, fig.height = 16, fig.height = 5}

awd_past_4weeks_map  + awd_past_week_map 

```

<!-- ## AJS map in the past week -->

```{r choropleth_ajs_pw_maps, message = FALSE, warning = FALSE}

## Join the past month dataset with map file
ajs_past_week <- map %>% 
  ## join the count data from past month verified dataset
  left_join(count_verified_signals_past_week(ebs_cleaned, "AJS"),
            by = "site_name") %>% 
    ## create levels for counts of ajs
   mutate(ajs_counts = fac_from_num(n)) %>% 
  ## remove the NA from ajs_counts
  filter(is.na(ajs_counts) == FALSE)



## Plot map of cases by camp
ajs_past_week_map <- ggplot() +
  ## add in the back ground tiles
  ggspatial::layer_spatial(tiles, interpolate = TRUE) +
  # shapefile as polygon
  geom_sf(data = ajs_past_week, aes(fill = ajs_counts)) + 
  # needed to avoid gridlines being drawn
  coord_sf(datum = NA) + 
  # add a scalebar
  annotation_scale() + 
  # choose palette colour for fill
  scale_fill_brewer(palette = "YlOrRd") +
  ## Add camp number as label
  geom_sf_text(data = ajs_past_week, aes(label = site_name), colour = "black",
               size = 2.5) +
  theme_void() +
  theme(plot.title = element_text(size= 12)) +
  labs(fill = "No. of signals",
       captions = str_glue("Source: MSF data from {reporting_week}"),
       title = str_glue("Number of verified AJS signals in \nCox's Bazar,{reporting_week}"))
  
```

<!-- ## AJS map in the past 4weeks -->

```{r choropleth_ajs_p4w_maps, message = FALSE, warning = FALSE}

## Join the past month dataset with map file
ajs_past_4weeks <- map %>% 
  ## join the count data from past month verified dataset
  left_join(count_verified_signals_past_4weeks(ebs_cleaned, "AJS"), 
            by = "site_name") %>% 
  ## create levels for counts of ajs
  mutate(ajs_counts = fac_from_num(n)) %>% 
  ## remove the empty values
  filter(is.na(ajs_counts) == FALSE)
         


## Plot map of cases by camp
ajs_past_4weeks_map <- ggplot() +
  ## add in the back ground tiles
  ggspatial::layer_spatial(tiles, interpolate = TRUE) +
  # shapefile as polygon
  geom_sf(data = ajs_past_4weeks, aes(fill = ajs_counts)) + 
  # needed to avoid gridlines being drawn
  coord_sf(datum = NA) + 
  # add a scalebar
  annotation_scale() + 
  # choose palette colour for fill
  scale_fill_brewer(palette = "YlOrRd") +
  ## Add camp number as label
  geom_sf_text(data = ajs_past_4weeks, aes(label = site_name), colour = "black",
               size = 2.5) +
  theme_void() +
  theme(plot.title = element_text(size= 12)) +
  labs(fill = "No. of signals",
       captions = str_glue("Source: MSF data from {three_weeks_ago} to {reporting_week}"),
       title = str_glue("Number of verified AJS signals in \nCox's Bazar, {three_weeks_ago}-{reporting_week}"))

```


<!-- #### Combine AJS maps -->
```{r combine_ajs_maps, fig.height = 16, fig.height = 5}

ajs_past_4weeks_map  + ajs_past_week_map 

```

<!-- ## Diphtheria map in the past week -->

```{r choropleth_diphtheria_pw_maps, message = FALSE, warning = FALSE}

## Join the past month dataset with map file
diphtheria_past_week <- map %>% 
  ## join the count data from past month verified dataset
  left_join(count_verified_signals_past_week(ebs_cleaned, "Diphtheria"),
            by = "site_name") %>% 
    ## create levels for counts of diphtheria
 mutate(diphtheria_counts = fac_from_num(n)) %>% 
  ## remove the NA from diphtheria_counts
  filter(is.na(diphtheria_counts) == FALSE)


## Plot map of cases by camp
diphtheria_past_week_map <- ggplot() +
  ## add in the back ground tiles
  ggspatial::layer_spatial(tiles, interpolate = TRUE) +
  # shapefile as polygon
  geom_sf(data = diphtheria_past_week, aes(fill = diphtheria_counts)) + 
  # needed to avoid gridlines being drawn
  coord_sf(datum = NA) + 
  # add a scalebar
  annotation_scale() + 
  # choose palette colour for fill
  scale_fill_brewer(palette = "YlOrRd") +
  ## Add camp number as label
  geom_sf_text(data = diphtheria_past_week, aes(label = site_name), colour = "black",
               size = 2.5) +
  theme_void() +
  theme(plot.title = element_text(size= 12)) +
  labs(fill = "No. of signals",
       captions = str_glue("Source: MSF data from {reporting_week}"),
       title = str_glue("Number of verified Diphtheria signals in \nCox's Bazar,{reporting_week}"))
  

```


<!-- ## diphtheria map in the past 4weeks -->

```{r choropleth_diphtheria_p4w_maps, message = FALSE, warning = FALSE}

## Join the past month dataset with map file
diphtheria_past_4weeks <- map %>% 
  ## join the count data from past month verified dataset
  left_join(count_verified_signals_past_4weeks(ebs_cleaned, "Diphtheria"), 
            by = "site_name") %>% 
  ## create levels for counts of diphtheria
  mutate(diphtheria_counts = fac_from_num(n)) %>% 
  ## remove the empty values
  filter(is.na(diphtheria_counts) == FALSE)
         


## Plot map of cases by camp
diphtheria_past_4weeks_map <- ggplot() +
  ## add in the back ground tiles
  ggspatial::layer_spatial(tiles, interpolate = TRUE) +
  # shapefile as polygon
  geom_sf(data = diphtheria_past_4weeks, aes(fill = diphtheria_counts)) + 
  # needed to avoid gridlines being drawn
  coord_sf(datum = NA) + 
  # add a scalebar
  annotation_scale() + 
  # choose palette colour for fill
  scale_fill_brewer(palette = "YlOrRd") +
  ## Add camp number as label
  geom_sf_text(data = diphtheria_past_4weeks, aes(label = site_name),
               colour = "black",
               size = 2.5) +
  theme_void() +
  theme(plot.title = element_text(size= 12)) +
  labs(fill = "No. of signals",
       captions = str_glue("Source: MSF data from {three_weeks_ago} to {reporting_week}"),
       title = str_glue("Number of verified Diphtheria signals in \nCox's Bazar, {three_weeks_ago}-{reporting_week}"))




```

<!-- #### Combine diphtheria maps -->
```{r combine_diphtheria_maps, fig.height = 16, fig.height = 5}

diphtheria_past_4weeks_map  + diphtheria_past_week_map 

```

<!-- ## Measles map in the past week -->

```{r choropleth_measles_pw_maps, message = FALSE, warning = FALSE}

## Join the past month dataset with map file
measles_past_week <- map %>%
  ## join the count data from past month verified dataset
  left_join(count_verified_signals_past_week(ebs_cleaned, "Measles"),
            by = "site_name") %>%
    ## create levels for counts of measles
 mutate(measles_counts = fac_from_num(n)) %>%
  ## remove the NA from measles_counts
  filter(is.na(measles_counts) == FALSE)


## Plot map of cases by camp
measles_past_week_map <- ggplot() +
  ## add in the back ground tiles
  ggspatial::layer_spatial(tiles, interpolate = TRUE) +
  # shapefile as polygon
  geom_sf(data = measles_past_week, aes(fill = measles_counts)) +
  # needed to avoid gridlines being drawn
  coord_sf(datum = NA) +
  # add a scalebar
  annotation_scale() +
  # choose palette colour for fill
  scale_fill_brewer(palette = "YlOrRd") +
  ## Add camp number as label
  geom_sf_text(data = measles_past_week, aes(label = site_name),
               colour = "black",
               size = 2.5) +
  theme_void() +
  theme(plot.title = element_text(size= 12)) +
  labs(fill = "No. of signals",
       captions = str_glue("Source: MSF data from {reporting_week}"),
       title = str_glue("Number of verified Measles signals in\n Cox's Bazar, {reporting_week}"))
  

```


<!-- ## measles map in the past 4weeks -->

```{r choropleth_measles_p4w_maps, message = FALSE, warning = FALSE}

# ## Join the past month dataset with map file
measles_past_4weeks <- map %>%
  ## join the count data from past month verified dataset
  left_join(count_verified_signals_past_4weeks(ebs_cleaned, "Measles"),
            by = "site_name") %>%
  ## create levels for counts of measles
  mutate(measles_counts = fac_from_num(n)) %>%
  ## remove the empty values
  filter(is.na(measles_counts) == FALSE)



## Plot map of cases by camp
measles_past_4weeks_map <- ggplot() +
  ## add in the back ground tiles
  ggspatial::layer_spatial(tiles, interpolate = TRUE) +
  # shapefile as polygon
  geom_sf(data = measles_past_4weeks, aes(fill = measles_counts)) +
  # needed to avoid gridlines being drawn
  coord_sf(datum = NA) +
  # add a scalebar
  annotation_scale() +
  # choose palette colour for fill
  scale_fill_brewer(palette = "YlOrRd") +
  ## Add camp number as label
  geom_sf_text(data = measles_past_4weeks, aes(label = site_name), colour = "black",
               size = 2.5) +
  theme_void() +
  theme(plot.title = element_text(size= 12)) +
  labs(fill = "No. of signals",
       captions = str_glue("Source: MSF data from {three_weeks_ago} to {reporting_week}"),
       title = str_glue("Number of verified measles signals in \nCox's Bazar, {three_weeks_ago}-{reporting_week}"))

```

<!-- #### Combine measles maps -->
```{r combine_measles_maps, fig.height = 16, fig.height = 5}

 measles_past_4weeks_map  + measles_past_week_map 

```


<!-- ## meningitis map in the past week -->

```{r choropleth_meningitis_pw_maps, message = FALSE, warning = FALSE}

## Join the past month dataset with map file
meningitis_past_week <- map %>%
  ## join the count data from past month verified dataset
  left_join(count_verified_signals_past_week(ebs_cleaned, "Meningitis"),
            by = "site_name") %>%
    ## create levels for counts of meningitis
 mutate(meningitis_counts = fac_from_num(n)) %>%
  ## remove the NA from meningitis_counts
  filter(is.na(meningitis_counts) == FALSE)


## Plot map of cases by camp
meningitis_past_week_map <- ggplot() +
  ## add in the back ground tiles
  ggspatial::layer_spatial(tiles, interpolate = TRUE) +
  # shapefile as polygon
  geom_sf(data = meningitis_past_week, aes(fill = meningitis_counts)) +
  # needed to avoid gridlines being drawn
  coord_sf(datum = NA) +
  # add a scalebar
  annotation_scale() +
  # choose palette colour for fill
  scale_fill_brewer(palette = "YlOrRd") +
  ## Add camp number as label
  geom_sf_text(data = meningitis_past_week, aes(label = site_name),
               colour = "black",
               size = 2.5) +
  theme_void() +
  theme(plot.title = element_text(size= 12)) +
  labs(fill = "No. of signals",
       captions = str_glue("Source: MSF data from {reporting_week}"),
       title = str_glue("Number of verified meningitis signals in\n Cox's Bazar, {reporting_week}"))
  

```


<!-- ## meningitis map in the past 4weeks -->

```{r choropleth_meningitis_p4w_maps, message = FALSE, warning = FALSE}

# ## Join the past month dataset with map file
meningitis_past_4weeks <- map %>%
  ## join the count data from past month verified dataset
  left_join(count_verified_signals_past_4weeks(ebs_cleaned, "Meningitis"),
            by = "site_name") %>%
  ## create levels for counts of meningitis
  mutate(meningitis_counts = fac_from_num(n)) %>%
  ## remove the empty values
  filter(is.na(meningitis_counts) == FALSE)



## Plot map of cases by camp
meningitis_past_4weeks_map <- ggplot() +
  ## add in the back ground tiles
  ggspatial::layer_spatial(tiles, interpolate = TRUE) +
  # shapefile as polygon
  geom_sf(data = meningitis_past_4weeks, aes(fill = meningitis_counts)) +
  # needed to avoid gridlines being drawn
  coord_sf(datum = NA) +
  # add a scalebar
  annotation_scale() +
  # choose palette colour for fill
  scale_fill_brewer(palette = "YlOrRd") +
  ## Add camp number as label
  geom_sf_text(data = meningitis_past_4weeks, aes(label = site_name), colour = "black",
               size = 2.5) +
  theme_void() +
  theme(plot.title = element_text(size= 12)) +
  labs(fill = "No. of signals",
       captions = str_glue("Source: MSF data from {three_weeks_ago} to {reporting_week}"),
       title = str_glue("Number of verified meningitis signals in \nCox's Bazar, {three_weeks_ago}-{reporting_week}"))

```

<!-- #### Combine meningitis maps -->
```{r combine_meningitis_maps, fig.height = 16, fig.height = 5}

meningitis_past_4weeks_map  + meningitis_past_week_map 

```



\newpage




# Annex 1: EBS definitions and responsibilities

**Signal**  A signal is reported data or information that represents a potential acute risk to human health. It is transmitted immediately and has not yet been verified as to whether or not it meets the case or event definition of the surveillance system. Signals are reported by CHVs, WHPs and TBAs.
**Verified signal**: Reported information that meets the formal signal definition. The EAT members verify the signals the day after they are reported. The epidemiologist reviews all verified signals to decide whether a risk assessment is required.
**Alert**: An alert will refer to a public health event that has been i) verified and ii) risk assessed and iii) requires an intervention (an investigation, a response or a communication)
**Response** A rapid response team is deployed and could conduct activities such as active case finding, vaccination, health promotion, WASH-related activities and case management  

  i)	Where teams are mobilized for the purpose of response against their routine activities  
  
  ii) New resources are mobilized or materials prepared
  
  iii) Communication done specifically for the alert

Response could be through:  (1) Routine epi teams, (2) Health promotion teams with special focus on the specific alert (3) Epi, WatSan and HP teams for alerts that require a combined response

HAO and Advocacy teams for alerts that need external communication


# Annex 2: Proposed threshold definitions
In order to make the EBS system more useful and relevant for the project, it is necessary to have a simple flagging system to be able to alert other team members of significant information. In the EBS SOP, there was a plan for risk assessments (to categorise alerts between minimal to severe risk) of verified signals by an assessment team in discussion with MTL/Medco/WASHCO- this step as well as the subsequent RRT deployment steps are currently missing. In part, this is due to the sheer volume of verified signals in this setting. In weeks 25- 34 we had 216 verified signals that EAT members reported required a response but no other actors were currently responding to. Typical EBS risk assessment steps are better suited to ad hoc early warning reports of acute public health events. With this in mind, we are proposing an intermediary step to filter routine alerts from unusual alerts that are more likely to require a risk assessment in the camps. However, given that other examples of how to do this in an MSF setting arent available, any feedback is very welcome! The below thresholds are a proposal for a way to filter alerts that would be indicative of geographically clustered transmission at camp-level, increasing transmission or high numbers of people affected- beyond what is typical for the setting. 

<!-- ## Table 1 -->
<!-- Create summary table of total verified signals by camp -->
## Threshold A: Most number of clusters per event type and camp in reporting week
The camp with the highest number of clusters per event type in `r reporting_week` is camp `r thresholda %>% pull(chv_signals_camp)` with `r thresholda %>% pull(n)` clusters of `r thresholda %>% pull(total_event_detail)` 

```{r threshold_A, fig.height = 3}
#Select highest/most concerning given context and describe whether risk assessment is required in the recommendation section

ebs_verified %>% 
  ## only keep signals reported during the reporting week 
  filter(epiweek_signal == reporting_week) %>% 
  ## only keep signals that require a response
  filter(verification_do_you_think_response_is_requi == "yes")  %>%
  ## only keep signals where no other actor is responding
  filter(verification_is_there_any_other_organizatio == "no")  %>%
  ## select variables to use for table
  select("Signal type"= total_event_detail, 
         chv_signals_camp) %>% 
  ## use tbl_summary from gtsummary package to make table
  tbl_summary(by = chv_signals_camp, 
              missing = "no")  %>% 
  bold_labels() %>%
  as_gt() %>%
  gt::gtsave(filename="thresholda.png", path = here::here("2021", "4_output", "archive"))

```


## Threshold B: Largest clusters
The largest single cluster in `r reporting_week` was for `r thresholdb %>% pull(total_event_detail)` in camp `r thresholdb %>% pull(chv_signals_camp)` that affected `r thresholdb %>% pull(verification_how_many_people)` people.

```{r threshold_B}
ebs_verified %>%
  ## only keep signals reported during the reporting week 
  filter(epiweek_signal == reporting_week) %>% 
  ## only keep signals that require a response
  filter(verification_do_you_think_response_is_requi == "yes")  %>%
  ## only keep signals where no other actor is responding
  filter(verification_is_there_any_other_organizatio == "no")  %>%
  group_by(chv_signals_camp) %>%
  filter(verification_how_many_people == max(verification_how_many_people) ) %>% 
              select("Camp"= chv_signals_camp,"signal type"= total_event_detail,  verification_how_many_people) %>%
  arrange(desc( verification_how_many_people)) %>%
  rename("Largest cluster- number of people"= verification_how_many_people) %>%
  kable()

```

## Threshold C: Increase in number of cases over time
There was a consecutive increase in number of people affected by `r thresholdc_increase %>% pull(Signals)` in the catchment area in the past 3 weeks.

```{r threshold_C}
kable(thresholdc)
```
